---
title: "Color Pallete examples based on DSMI"
author: "Charlie Whittaker"
output: 
  html_document:
    toc: true
    toc_depth: 4
  html_notebook: default
date: "3/3/2026"
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = TRUE,
  warning = TRUE,
  error = TRUE,
  cache = TRUE,
  cache.lazy = FALSE,
  fig.retina = 2
)

# For bioinformatics (Seurat, parallel)
options(future.globals.maxSize = 8e10)

# Optional: seed for reproducibility
set.seed(42)

# Clear knitr cache
unlink("hts_Processing_cache", recursive = TRUE)
```

```{r}
library(openxlsx)
library(tidyverse)
library(ggrepel)
library(ggpubr)
library(purrr)
library(knitr)
library(lubridate)
library(patchwork)
```

# Data Import

```{r}
dsmi.final <- read.xlsx("dsmi.final.xlsx")
```

# Training Data

## Extract the data from the main dataframe

```{r}
train <- dsmi.final %>% dplyr::filter(Category == "Training")
dim(train)
old.train <- dsmi.final %>% dplyr::filter(Category == "Hours of Service " & Charge.Name == "Training Session")
dim(old.train)
train <- rbind(train,old.train)
dim(train)
```

# Color Pallete examples

## Current example

```{r}
# Define colors for each Sponsor (adjust these hex codes as needed)
sponsor_colors <- c(
  "Center_Member" = "steelblue2",
  "MIT_Sponsor" = "orange",
  "Non_Sponsor" = "#C16644"
)

# Define order of sponsors for stacking (top to bottom)
sponsor_order <- c("Center_Member", "MIT_Sponsor", "Non_Sponsor")
```

### Current colors pie chart

```{r}
current.pie <- train %>%
  mutate(Sponsor_ForPlot = factor(Sponsor_ForPlot, levels = sponsor_order)) %>%
  group_by(Sponsor_ForPlot) %>%
  summarise(Quantity = sum(round(Quantity,0)), 
            Num_Labs = if_else(first(Sponsor_ForPlot) == "Center_Member",
                               n_distinct(Faculty.Key[IGT_DSMI_Sponsor == "KI_member"]),
                               n_distinct(Faculty.Key)),
            .groups = "drop") %>%
  mutate(Percent = round((Quantity / sum(Quantity)) * 100, 0),
         Lab_Label = if_else(Sponsor_ForPlot == "Center_Member",
                             paste0(Num_Labs, " Members"),
                             paste0(Num_Labs, " Users")),
         Label = paste0(Quantity, "\n(", Percent, "%)\n", Lab_Label)) %>%
  ggplot(aes(x = "", y = Quantity, fill = Sponsor_ForPlot)) +
  geom_col() +
  geom_text(aes(label = Label), 
            position = position_stack(vjust = 0.5),
            size = 3.6,
            fontface = "bold",
            color = "black") +
  coord_polar("y", start = 0) +
  scale_fill_manual(values = sponsor_colors) +
  labs(title = "current\n(attendees)",
       fill = "Sponsor") +
  theme_void() +
  theme(legend.position = "bottom",
        plot.title = element_text(size = 12,
                                  face = "plain",
                                  hjust = 0.5))

current.pie
```

## sarah picks

```{r}
# Define colors for each Sponsor (adjust these hex codes as needed)
sponsor_colors <- c(
  "Center_Member" = "#CB674C",
  "MIT_Sponsor" = "#6C95B0", 
  "Non_Sponsor" = "#DBD7D3"
)

# Define order of sponsors for stacking (top to bottom)
sponsor_order <- c("Center_Member", "MIT_Sponsor", "Non_Sponsor")
```

### sarah picks pie

```{r}
sarah.pie <- train %>%
  mutate(Sponsor_ForPlot = factor(Sponsor_ForPlot, levels = sponsor_order)) %>%
  group_by(Sponsor_ForPlot) %>%
  summarise(Quantity = sum(round(Quantity,0)), 
            Num_Labs = if_else(first(Sponsor_ForPlot) == "Center_Member",
                               n_distinct(Faculty.Key[IGT_DSMI_Sponsor == "KI_member"]),
                               n_distinct(Faculty.Key)),
            .groups = "drop") %>%
  mutate(Percent = round((Quantity / sum(Quantity)) * 100, 0),
         Lab_Label = if_else(Sponsor_ForPlot == "Center_Member",
                             paste0(Num_Labs, " Members"),
                             paste0(Num_Labs, " Users")),
         Label = paste0(Quantity, "\n(", Percent, "%)\n", Lab_Label)) %>%
  ggplot(aes(x = "", y = Quantity, fill = Sponsor_ForPlot)) +
  geom_col() +
  geom_text(aes(label = Label), 
            position = position_stack(vjust = 0.5),
            size = 3.6,
            fontface = "bold",
            color = "black") +
  coord_polar("y", start = 0) +
  scale_fill_manual(values = sponsor_colors) +
  labs(title = "Sarah Picks\n(attendees)",
       fill = "Sponsor") +
  theme_void() +
  theme(legend.position = "bottom",
        plot.title = element_text(size = 12,
                                  face = "plain",
                                  hjust = 0.5))

sarah.pie
```

## colorblind safe example 1

```{r}
# Define colors for each Sponsor (adjust these hex codes as needed)
sponsor_colors <- c(
  "Center_Member" = "#0173B2",
  "MIT_Sponsor" = "#DE8F05", 
  "Non_Sponsor" = "#029E73"
)

# Define order of sponsors for stacking (top to bottom)
sponsor_order <- c("Center_Member", "MIT_Sponsor", "Non_Sponsor")
```

### colorblind safe example 1

```{r}
cb1.pie <- train %>%
  mutate(Sponsor_ForPlot = factor(Sponsor_ForPlot, levels = sponsor_order)) %>%
  group_by(Sponsor_ForPlot) %>%
  summarise(Quantity = sum(round(Quantity,0)), 
            Num_Labs = if_else(first(Sponsor_ForPlot) == "Center_Member",
                               n_distinct(Faculty.Key[IGT_DSMI_Sponsor == "KI_member"]),
                               n_distinct(Faculty.Key)),
            .groups = "drop") %>%
  mutate(Percent = round((Quantity / sum(Quantity)) * 100, 0),
         Lab_Label = if_else(Sponsor_ForPlot == "Center_Member",
                             paste0(Num_Labs, " Members"),
                             paste0(Num_Labs, " Users")),
         Label = paste0(Quantity, "\n(", Percent, "%)\n", Lab_Label)) %>%
  ggplot(aes(x = "", y = Quantity, fill = Sponsor_ForPlot)) +
  geom_col() +
  geom_text(aes(label = Label), 
            position = position_stack(vjust = 0.5),
            size = 3.6,
            fontface = "bold",
            color = "black") +
  coord_polar("y", start = 0) +
  scale_fill_manual(values = sponsor_colors) +
  labs(title = "CBsafe 1\n(attendees)",
       fill = "Sponsor") +
  theme_void() +
  theme(legend.position = "bottom",
        plot.title = element_text(size = 12,
                                  face = "plain",
                                  hjust = 0.5))

cb1.pie
```

## colorblind safe example 2

```{r}
# Define colors for each Sponsor (adjust these hex codes as needed)
sponsor_colors <- c(
  "Center_Member" = "#0173B2",
  "MIT_Sponsor" = "#DE8F05", 
  "Non_Sponsor" = "#CC78BC"
)

# Define order of sponsors for stacking (top to bottom)
sponsor_order <- c("Center_Member", "MIT_Sponsor", "Non_Sponsor")
```

### colorblind safe example 2

```{r}
cb2.pie <- train %>%
  mutate(Sponsor_ForPlot = factor(Sponsor_ForPlot, levels = sponsor_order)) %>%
  group_by(Sponsor_ForPlot) %>%
  summarise(Quantity = sum(round(Quantity,0)), 
            Num_Labs = if_else(first(Sponsor_ForPlot) == "Center_Member",
                               n_distinct(Faculty.Key[IGT_DSMI_Sponsor == "KI_member"]),
                               n_distinct(Faculty.Key)),
            .groups = "drop") %>%
  mutate(Percent = round((Quantity / sum(Quantity)) * 100, 0),
         Lab_Label = if_else(Sponsor_ForPlot == "Center_Member",
                             paste0(Num_Labs, " Members"),
                             paste0(Num_Labs, " Users")),
         Label = paste0(Quantity, "\n(", Percent, "%)\n", Lab_Label)) %>%
  ggplot(aes(x = "", y = Quantity, fill = Sponsor_ForPlot)) +
  geom_col() +
  geom_text(aes(label = Label), 
            position = position_stack(vjust = 0.5),
            size = 3.6,
            fontface = "bold",
            color = "black") +
  coord_polar("y", start = 0) +
  scale_fill_manual(values = sponsor_colors) +
  labs(title = "CBsafe 2\n(attendees)",
       fill = "Sponsor") +
  theme_void() +
  theme(legend.position = "bottom",
        plot.title = element_text(size = 12,
                                  face = "plain",
                                  hjust = 0.5))

cb2.pie
```

## colorblind safe example 3

```{r}
# Define colors for each Sponsor (adjust these hex codes as needed)
sponsor_colors <- c(
  "Center_Member" = "#56B4E9",
  "MIT_Sponsor" = "#E69F00", 
  "Non_Sponsor" = "#009E73"
)

# Define order of sponsors for stacking (top to bottom)
sponsor_order <- c("Center_Member", "MIT_Sponsor", "Non_Sponsor")
```

### colorblind safe example 3

```{r}
cb3.pie <- train %>%
  mutate(Sponsor_ForPlot = factor(Sponsor_ForPlot, levels = sponsor_order)) %>%
  group_by(Sponsor_ForPlot) %>%
  summarise(Quantity = sum(round(Quantity,0)), 
            Num_Labs = if_else(first(Sponsor_ForPlot) == "Center_Member",
                               n_distinct(Faculty.Key[IGT_DSMI_Sponsor == "KI_member"]),
                               n_distinct(Faculty.Key)),
            .groups = "drop") %>%
  mutate(Percent = round((Quantity / sum(Quantity)) * 100, 0),
         Lab_Label = if_else(Sponsor_ForPlot == "Center_Member",
                             paste0(Num_Labs, " Members"),
                             paste0(Num_Labs, " Users")),
         Label = paste0(Quantity, "\n(", Percent, "%)\n", Lab_Label)) %>%
  ggplot(aes(x = "", y = Quantity, fill = Sponsor_ForPlot)) +
  geom_col() +
  geom_text(aes(label = Label), 
            position = position_stack(vjust = 0.5),
            size = 3.6,
            fontface = "bold",
            color = "black") +
  coord_polar("y", start = 0) +
  scale_fill_manual(values = sponsor_colors) +
  labs(title = "CBsafe 3\n(gold standard according to LLM)",
       fill = "Sponsor") +
  theme_void() +
  theme(legend.position = "bottom",
        plot.title = element_text(size = 12,
                                  face = "plain",
                                  hjust = 0.5))

cb3.pie
```

## colorblind safe example 4

```{r}
# Define colors for each Sponsor (adjust these hex codes as needed)
sponsor_colors <- c(
  "Center_Member" = "#66C2A5",
  "MIT_Sponsor" = "#FC8D62", 
  "Non_Sponsor" = "#8DA0CB"
)

# Define order of sponsors for stacking (top to bottom)
sponsor_order <- c("Center_Member", "MIT_Sponsor", "Non_Sponsor")
```

### colorblind safe example 4

```{r}
cb4.pie <- train %>%
  mutate(Sponsor_ForPlot = factor(Sponsor_ForPlot, levels = sponsor_order)) %>%
  group_by(Sponsor_ForPlot) %>%
  summarise(Quantity = sum(round(Quantity,0)), 
            Num_Labs = if_else(first(Sponsor_ForPlot) == "Center_Member",
                               n_distinct(Faculty.Key[IGT_DSMI_Sponsor == "KI_member"]),
                               n_distinct(Faculty.Key)),
            .groups = "drop") %>%
  mutate(Percent = round((Quantity / sum(Quantity)) * 100, 0),
         Lab_Label = if_else(Sponsor_ForPlot == "Center_Member",
                             paste0(Num_Labs, " Members"),
                             paste0(Num_Labs, " Users")),
         Label = paste0(Quantity, "\n(", Percent, "%)\n", Lab_Label)) %>%
  ggplot(aes(x = "", y = Quantity, fill = Sponsor_ForPlot)) +
  geom_col() +
  geom_text(aes(label = Label), 
            position = position_stack(vjust = 0.5),
            size = 3.6,
            fontface = "bold",
            color = "black") +
  coord_polar("y", start = 0) +
  scale_fill_manual(values = sponsor_colors) +
  labs(title = "CBsafe 4\n(attendees)",
       fill = "Sponsor") +
  theme_void() +
  theme(legend.position = "bottom",
        plot.title = element_text(size = 12,
                                  face = "plain",
                                  hjust = 0.5))

cb4.pie
```

# Plot the pie charts figure

```{r,fig.width=8,fig.height=12}
current.pie + sarah.pie + cb1.pie + cb2.pie + cb3.pie + cb4.pie + plot_layout(ncol = 2, guides = "collect") & 
  theme(legend.position = "bottom")


combo.pie <- current.pie + sarah.pie + cb1.pie + cb2.pie + cb3.pie + cb4.pie + plot_layout(ncol = 2, guides = "collect") & 
  theme(legend.position = "bottom")

ggsave("figures/ColorTest.pie.pdf", 
       plot = combo.pie, 
       dpi = 600, 
       width = 8, 
       height = 12, 
       units = "in")
```

# write session info

Capturing information about the R session helps document your work

```{r}
sessionInfo()
writeLines(capture.output(sessionInfo()), "Color_Tests_sessionInfo.txt")
```
