---
title: "Flow - Data Cleaning, Summary, and Visualizations"
author: "Charlie Whittaker"
output: 
  html_document:
    toc: true
    toc_depth: 4
  html_notebook: default
date: "2/15/2026"
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = TRUE,
  warning = TRUE,
  error = TRUE,
  cache = TRUE,
  cache.lazy = FALSE,
  fig.retina = 2
)

# For bioinformatics (Seurat, parallel)
options(future.globals.maxSize = 8e10)

# Optional: seed for reproducibility
set.seed(42)

# Clear knitr cache
unlink("PMIT_Processing_cache", recursive = TRUE)
```

```{r}
library(openxlsx)
library(tidyverse)
library(ggrepel)
library(ggpubr)
library(purrr)
library(knitr)
library(lubridate)
library(patchwork)
```

# Data Import

## Faculty/Lab Annotations using Faculty.key

A reference file has been generated that associates faculty and/or lab values that appear in all accounting data with a harmonized faculty key
This file is a work in progress

### Import faculty data

```{r}
fac.key <- read.xlsx("../data/FacultyKeys.xlsx")
```

### Read in ilabs stock data and select Flow Cytometry Core rows

```{r}
ilabs <- read.xlsx("../data/stock_ilabs.xlsx")
ilabs.flow <- ilabs %>% dplyr::filter(Core.Name == "Flow Cytometry Core")
rm(ilabs)
```

### Exclude Nanowell services 

The nanowell core services are included in Flow Cytometry ilabs data and need to be excluded - they will be reported on in the HTS section

```{r}
cat("Dimensions of ilabs.flow before exlusion of nanowell:", dim(ilabs.flow), "\n")
ilabs.flow <- ilabs.flow %>% dplyr::filter(Category %in% c("Analysis Machines", "Cell Sorting Equipment"))
cat("Dimensions of ilabs.flow after exlusion of nanowell:", dim(ilabs.flow), "\n")
```

### Repair date

```{r}
ilabs.flow$Purchase.Date <- as.Date(ilabs.flow$Purchase.Date, origin = "1899-12-30")
```

### Check dates before reporting range selection

```{r, fig.width=12,fig.height=5}
ilabs.flow %>%
  mutate(Year_Month = floor_date(Purchase.Date, "month")) %>%
  count(Year_Month) %>%
  arrange(Year_Month) %>%
  ggplot(aes(x = Year_Month, y = n)) +
  geom_col(fill = "steelblue") +
  scale_x_date(date_labels = "%b %Y", date_breaks = "1 month") +
  labs(title = "Records by Month",
       x = "Month",
       y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Select reporting date range

```{r}
cat("Dimensions of ilabs.flow before reporting range selection:", dim(ilabs.flow), "\n")
ilabs.flow <- ilabs.flow %>% dplyr::filter(between(Purchase.Date, as.Date("2020-05-01"), as.Date("2025-10-31")))
cat("Dimensions of ilabs.flow after reporting range selection:", dim(ilabs.flow), "\n")
```

### Check dates after reporting range selection

```{r, fig.width=12,fig.height=5}
ilabs.flow %>%
  mutate(Year_Month = floor_date(Purchase.Date, "month")) %>%
  count(Year_Month) %>%
  arrange(Year_Month) %>%
  ggplot(aes(x = Year_Month, y = n)) +
  geom_col(fill = "steelblue") +
  scale_x_date(date_labels = "%b %Y", date_breaks = "1 month") +
  labs(title = "Records by Month",
       x = "Month",
       y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

# Service Annotation

## Simplify Charge.Name

The Charge.Name data is "|" separated values, the first value is most useful, extract it and store in a new column

```{r}
ilabs.flow <- ilabs.flow %>% mutate(Charge.Name.Simple = str_trim(str_extract(Charge.Name, "^[^|]*")))
```

## Examine Service.Type column

One spurious nanowell value remained - exclude it here

```{r}
table(ilabs.flow$Service.Type)
ilabs.flow <- ilabs.flow %>% dplyr::filter(Service.Type != "Nanowell cytometry device")
table(ilabs.flow$Service.Type)
```

## Examine the data

```{r}
table(ilabs.flow$Category)
# table(ilabs.flow$Charge.Name.Simple)
```

### Update category with Imaging Cytometry

```{r}
ilabs.flow <- ilabs.flow %>%
  mutate(Category = if_else(str_starts(Charge.Name.Simple, "Amnis"), 
                             "Imaging Cytometry", 
                             Category))
```

# Faculty Annotation Work

## Associate Faculty.Key with ilabs data

```{r}
ilabs.flow <- ilabs.flow %>%
  left_join(fac.key %>%
              select(ValueInData, Faculty.Key) %>%
              distinct(), 
            by = c("Customer.Lab" = "ValueInData"))
```

## Import lab annotation data

NOTE: Once flow sponsor groups are finalized, remove the tmp designation

```{r}
fac.annot <- read.xlsx("../data/annotated.Faculty.Keys.xlsx")
```

### Join to ilabs.flow using Faculty.Key

```{r}
ilabs.flow <- ilabs.flow %>%
  dplyr::left_join(
    fac.annot %>% dplyr::select(Faculty.Key, Flow_Sponsor),
    by = "Faculty.Key"
  )
```

## Simplify sponsor associations for plotting

### Import simplified categories

```{r}
sponsor.toPlot <- read.xlsx("../data/sponsor.Keys.ForPlotting.xlsx")
```

```{r}
sponsor.toPlot
```

### Join to ilabs.flow using

```{r}
ilabs.flow <- ilabs.flow %>%
  dplyr::left_join(sponsor.toPlot, by = c("Flow_Sponsor" = "Sponsor"))
```

# Data Visualizations

## Define Color Palette

```{r}
# Define colors for each Sponsor (adjust these hex codes as needed)
# Original palette
# sponsor_colors <- c(
#   "Center_Member" = "steelblue2",
#   "MIT_Sponsor" = "orange",
#   "Non_Sponsor" = "#C16644"
# )

# Colorblind safe 3
sponsor_colors <- c(
  "Center_Member" = "#56B4E9",
  "MIT_Sponsor" = "#E69F00",
  "Non_Sponsor" = "#009E73"
)

# Define order of sponsors for stacking (top to bottom)
sponsor_order <- c("Center_Member", "MIT_Sponsor", "Non_Sponsor")
```

## Instrument Use: Training (hours)
1. Usage.Type contains “Training” or “training” (n=1887)

### Select the data

```{r}
train <- ilabs.flow %>%dplyr::filter(str_detect(Usage.Type, "Training|training"))
```

### Prepare Instrument Use: Training (hours) pie chart

```{r}
train.pie <- train %>%
  mutate(Sponsor_ForPlot = factor(Sponsor_ForPlot, levels = sponsor_order)) %>%
  group_by(Sponsor_ForPlot) %>%
  summarise(Quantity = sum(round(Quantity,0)), 
            Num_Labs = if_else(first(Sponsor_ForPlot) == "Center_Member",
                               n_distinct(Faculty.Key[Flow_Sponsor == "KI_member"]),
                               n_distinct(Faculty.Key)),
            .groups = "drop") %>%
  mutate(Percent = round((Quantity / sum(Quantity)) * 100, 0),
         Lab_Label = if_else(Sponsor_ForPlot == "Center_Member",
                             paste0(Num_Labs, " Members"),
                             paste0(Num_Labs, " Users")),
         Label = paste0(Quantity, "\n", Percent, "%\n", Lab_Label)) %>%
  ggplot(aes(x = "", y = Quantity, fill = Sponsor_ForPlot)) +
  geom_col() +
  geom_text(aes(label = Label), 
            position = position_stack(vjust = 0.5),
            size = 3.6,
            fontface = "bold",
            color = "black") +
  coord_polar("y", start = 0) +
  scale_fill_manual(values = sponsor_colors) +
  labs(title = "Training (hours)",
       fill = "Sponsor") +
  theme_void() +
  theme(legend.position = "bottom",
        plot.title = element_text(size = 12,
                                  face = "plain",
                                  hjust = 0.5))

train.pie
```

## Instrument Use: Trained- Sorting (hours)
1. Usage.Type does not contain “Training” or “training”
1. Category = “Cell Sorting Equipment" (n=12194)

### Select the data

```{r}
sort <- ilabs.flow %>% dplyr::filter(Category == "Cell Sorting Equipment") %>%
  dplyr::filter(!str_detect(Usage.Type, "Training|training"))
```

### Prepare Instrument Use: Trained- Sorting (hours) pie chart

```{r}
sort.pie <- sort %>%
  mutate(Sponsor_ForPlot = factor(Sponsor_ForPlot, levels = sponsor_order)) %>%
  group_by(Sponsor_ForPlot) %>%
  summarise(Quantity = sum(round(Quantity,0)), 
            Num_Labs = if_else(first(Sponsor_ForPlot) == "Center_Member",
                               n_distinct(Faculty.Key[Flow_Sponsor == "KI_member"]),
                               n_distinct(Faculty.Key)),
            .groups = "drop") %>%
  mutate(Percent = round((Quantity / sum(Quantity)) * 100, 0),
         Lab_Label = if_else(Sponsor_ForPlot == "Center_Member",
                             paste0(Num_Labs, " Members"),
                             paste0(Num_Labs, " Users")),
         Label = paste0(Quantity, "\n", Percent, "%\n", Lab_Label)) %>%
  ggplot(aes(x = "", y = Quantity, fill = Sponsor_ForPlot)) +
  geom_col() +
  geom_text(aes(label = Label), 
            position = position_stack(vjust = 0.5),
            size = 3.6,
            fontface = "bold",
            color = "black") +
  coord_polar("y", start = 0) +
  scale_fill_manual(values = sponsor_colors) +
  labs(title = "Sorting (hours)",
       fill = "Sponsor") +
  theme_void() +
  theme(legend.position = "bottom",
        plot.title = element_text(size = 12,
                                  face = "plain",
                                  hjust = 0.5))

sort.pie
```

## Instrument Use: Trained- Analysis (hours)
1. Usage.Type does not contain “Training” or “training”
1. Category = “Analysis Machines" (n=37338)

### Select the data

```{r}
trained.analysis <- ilabs.flow %>% dplyr::filter(Category == "Analysis Machines") %>%
  dplyr::filter(!str_detect(Usage.Type, "Training|training"))
```

### Prepare Instrument Use: Trained- Sorting (hours) pie chart

```{r}
trained.analysis.pie <- trained.analysis %>%
  mutate(Sponsor_ForPlot = factor(Sponsor_ForPlot, levels = sponsor_order)) %>%
  group_by(Sponsor_ForPlot) %>%
  summarise(Quantity = sum(round(Quantity,0)), 
            Num_Labs = if_else(first(Sponsor_ForPlot) == "Center_Member",
                               n_distinct(Faculty.Key[Flow_Sponsor == "KI_member"]),
                               n_distinct(Faculty.Key)),
            .groups = "drop") %>%
  mutate(Percent = round((Quantity / sum(Quantity)) * 100, 0),
         Lab_Label = if_else(Sponsor_ForPlot == "Center_Member",
                             paste0(Num_Labs, " Members"),
                             paste0(Num_Labs, " Users")),
         Label = paste0(Quantity, "\n", Percent, "%\n", Lab_Label)) %>%
  ggplot(aes(x = "", y = Quantity, fill = Sponsor_ForPlot)) +
  geom_col() +
  geom_text(aes(label = Label), 
            position = position_stack(vjust = 0.5),
            size = 3.6,
            fontface = "bold",
            color = "black") +
  coord_polar("y", start = 0) +
  scale_fill_manual(values = sponsor_colors) +
  labs(title = "Trained - Sorting (hours)",
       fill = "Sponsor") +
  theme_void() +
  theme(legend.position = "bottom",
        plot.title = element_text(size = 12,
                                  face = "plain",
                                  hjust = 0.5))

trained.analysis.pie
```

## Instrument Use: Trained- Imaging Cytometry (hours)
1. Usage.Type does not contain “Training” or “training”
1. Category = “Imaging Cytometry" (n=290)

```{r}
cytometry <- ilabs.flow %>% dplyr::filter(Category == "Imaging Cytometry") %>%
  dplyr::filter(!str_detect(Usage.Type, "Training|training"))
```

### Prepare Instrument Use: Trained- Imaging Cytometry (hours) pie chart

```{r}
cytometry.pie <- cytometry %>%
  mutate(Sponsor_ForPlot = factor(Sponsor_ForPlot, levels = sponsor_order)) %>%
  group_by(Sponsor_ForPlot) %>%
  summarise(Quantity = sum(round(Quantity,0)), 
            Num_Labs = if_else(first(Sponsor_ForPlot) == "Center_Member",
                               n_distinct(Faculty.Key[Flow_Sponsor == "KI_member"]),
                               n_distinct(Faculty.Key)),
            .groups = "drop") %>%
  mutate(Percent = round((Quantity / sum(Quantity)) * 100, 0),
         Lab_Label = if_else(Sponsor_ForPlot == "Center_Member",
                             paste0(Num_Labs, " Members"),
                             paste0(Num_Labs, " Users")),
         Label = paste0(Quantity, "\n", Percent, "%\n", Lab_Label)) %>%
  ggplot(aes(x = "", y = Quantity, fill = Sponsor_ForPlot)) +
  geom_col() +
  geom_text(aes(label = Label), 
            position = position_stack(vjust = 0.5),
            size = 3.6,
            fontface = "bold",
            color = "black") +
  coord_polar("y", start = 0) +
  scale_fill_manual(values = sponsor_colors) +
  labs(title = "Trained - Imaging Cytometry (hours)",
       fill = "Sponsor") +
  theme_void() +
  theme(legend.position = "bottom",
        plot.title = element_text(size = 12,
                                  face = "plain",
                                  hjust = 0.5))

cytometry.pie
```

# Plot the pie charts figure

```{r,fig.width=14,fig.height=4}
train.pie + sort.pie + trained.analysis.pie + cytometry.pie + plot_layout(ncol = 4, guides = "collect") & 
  theme(legend.position = "bottom")

combo.pie <- train.pie + sort.pie + trained.analysis.pie + cytometry.pie + plot_layout(ncol = 4, guides = "collect") & 
  theme(legend.position = "bottom")
ggsave("figures/FLOW_combo.pie.pdf",
       plot = combo.pie,
       dpi = 600,
       width = 14,
       height = 4,
       units = "in")
```

# Write annotated output file

```{r}
write.xlsx(ilabs.flow,"flow.final.xlsx")
```

# write session info

Capturing information about the R session helps document your work

```{r}
sessionInfo()
writeLines(capture.output(sessionInfo()), "Flow_Processing_sessionInfo.txt")
```
