---
title: "Genomics External Data Prep"
author: "Charlie Whittaker"
output: 
  html_document:
    toc: true
    toc_depth: 4
  html_notebook: default
date: "12/15/2025"
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = TRUE,
  warning = TRUE,
  error = TRUE,
  cache = TRUE,
  cache.lazy = FALSE,
  fig.retina = 2,
  fig.width = 6,
  fig.height = 6,
  dpi = 300
)

# For bioinformatics (Seurat, parallel)
options(future.globals.maxSize = 8e10)

# Optional: seed for reproducibility
set.seed(42)
```

```{r}
library(openxlsx)
library(tidyverse)
library(ggrepel)
library(ggpubr)
library(purrr)
library(knitr)
library(lubridate)
```

# Data Import

# Read in BMC External data

```{r}
bmc.ext <- read.xlsx("../data/ExtUsage2025_2510.xlsx")
```

# Data Repair - BMC file

## exclude infomatics data from genomics file

```{r}
bmc.ext.filt <- bmc.ext %>% dplyr::filter(TYPE != "Bioinformatics")
```

# Export BMC LAB data

```{r}
BMC.LAB.data <- bmc.ext.filt %>% dplyr::select(LAB)
```

```{r}
BMC.LAB.data.distinct <- BMC.LAB.data %>% distinct() 
```

```{r}
write.xlsx(BMC.LAB.data.distinct,"../data/BMC.LAB.data.xlsx")
```

## Repair the date

The imported dates are in a machine-readable format that is not
valuable. convert to human readable:

```{r}
bmc.ext.filt$Purchase.Date <- as.Date(bmc.ext.filt$DATE, origin = "1899-12-30")
```

remove the old date and other unnecessary columns

```{r}
bmc.ext.filt <- bmc.ext.filt %>% dplyr::select(-DATE,-CostObj,-TargetObject,-`_ID`,-PROJID)
```

Rename columns to match ilabs

```{r}
bmc.ext.filt <- bmc.ext.filt %>% dplyr::rename(Customer.Lab = LAB)
bmc.ext.filt <- bmc.ext.filt %>% dplyr::rename(Customer.Name = USER)
bmc.ext.filt <- bmc.ext.filt %>% dplyr::rename(Service.Type = DETAIL)
bmc.ext.filt <- bmc.ext.filt %>% dplyr::rename(Quantity = NUM)
bmc.ext.filt <- bmc.ext.filt %>% dplyr::rename(Total.Price = COST)
bmc.ext.filt <- bmc.ext.filt %>% dplyr::rename(Category = TYPE)
```

Add missing columns needed to match ilabs data

```{r}
bmc.ext.filt <- bmc.ext.filt %>% dplyr::mutate(Service.ID = "FilemakerData")
bmc.ext.filt <- bmc.ext.filt %>% dplyr::mutate(Customer.Institute = "")
bmc.ext.filt <- bmc.ext.filt %>% dplyr::mutate(Customer.Department = "")
bmc.ext.filt <- bmc.ext.filt %>% dplyr::mutate(Charge.Name = Service.Type)
bmc.ext.filt <- bmc.ext.filt %>% dplyr::mutate(Unit.of.Measure = "")
bmc.ext.filt <- bmc.ext.filt %>% dplyr::mutate(Notes = "")
bmc.ext.filt <- bmc.ext.filt %>% dplyr::mutate(Price = "")
bmc.ext.filt <- bmc.ext.filt %>% dplyr::mutate(Price.Type = "External non-profit")
bmc.ext.filt <- bmc.ext.filt %>% dplyr::mutate(Center = "MIT-BMC-CORE")
bmc.ext.filt <- bmc.ext.filt %>% dplyr::mutate(Fac.Group = "MIT_non.sponsor.external")
bmc.ext.filt <- bmc.ext.filt %>% dplyr::mutate(Core.Name = "KI Genomics Core / MIT BioMicro Center")
bmc.ext.filt <- bmc.ext.filt %>% dplyr::mutate(Core.ID = "")
bmc.ext.filt <- bmc.ext.filt %>% dplyr::mutate(temp.Inst = "")
bmc.ext.filt <- bmc.ext.filt %>% dplyr::mutate(Usage.Type = "")
bmc.ext.filt <- bmc.ext.filt %>% dplyr::mutate(Customer.Department = "")
```

Reorder columns to match ilabs

```{r}
bmc.ext.filt <- bmc.ext.filt %>% dplyr::select(Customer.Lab,Service.ID,Service.Type,Customer.Name,Customer.Department,Customer.Institute,Charge.Name,
                                               Notes,Unit.of.Measure,Quantity,Price,Total.Price,Price.Type,Purchase.Date,Core.Name,Center,
                                               Category,Usage.Type,Core.ID,temp.Inst,Fac.Group)
```

# Join ilabs and filemaker

```{r}
ki.ilabs.genomics.plusFilemaker <- rbind(ki.ilabs.simple.final.genomics,bmc.ext.filt)
```

```{r}
write.xlsx(ki.ilabs.genomics.plusFilemaker,"Parsed_ki.ilabs.genomics.plusFilemaker.xlsx")
```


<!-- ## Repairing Service.Type -->

<!-- There are some storage entries in the dataframe: -->

<!-- ```{r} -->
<!-- length(unique(ki.ilabs.simple$Service.Type)) -->
<!-- ``` -->

<!-- Exclude these charges with the following command, note, this same -->
<!-- approach, and its inverse, could be applied to the entire ilabs export -->
<!-- It was manually done on the input data -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple <- ilabs.bioinfo.simple %>%  -->
<!--   dplyr::filter(!Service.Type %in% c("Data Storage Charge", "Data Storage Monthly Charge")) -->
<!-- dim(ilabs.bioinfo.simple) -->
<!-- ``` -->

<!-- The column Service.Type has some overly complicated values and some -->
<!-- muddled information. these commands will make repairs: -->

<!-- any time the string "Training" appears in a value, the value should be -->
<!-- renamed to "Training_Session" -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple <- ilabs.bioinfo.simple %>% -->
<!--   dplyr::mutate(Service.Type = case_when( -->
<!--     str_detect(Service.Type, regex("Training", ignore_case = TRUE)) ~ "Training_Session", -->
<!--     TRUE ~ Service.Type -->
<!--   )) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- table(ilabs.bioinfo.simple$Service.Type) -->
<!-- ``` -->

<!-- The rows annotated with Bioinformatics Project Consult are in error Some -->
<!-- should actually be Bioinformatics Project Request or Prepaid -->
<!-- Bioinformatics Support -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple %>% dplyr::filter(Service.Type == "Bioinformatics Project Consult") %>% dplyr::select(Service.ID) %>% table() -->
<!-- ``` -->

<!-- These mis-annotated rows are the only occurences of those service IDs in -->
<!-- the data -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple %>% dplyr::filter(Service.ID == "BCC-AMON-prepaid bioinformatics 0.4 - 2021")  %>% dplyr::select(Service.ID) %>% table() -->
<!-- ilabs.bioinfo.simple %>% dplyr::filter(Service.ID == "BCC-HS-2606")  %>% dplyr::select(Service.ID) %>% table() -->
<!-- ``` -->

<!-- Conditionally rename Service.Type for these ServiceIDs -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple <- ilabs.bioinfo.simple %>% -->
<!--   mutate(Service.Type = case_when( -->
<!--     Service.ID == "BCC-AMON-prepaid bioinformatics 0.4 - 2021" ~ "Prepaid Bioinformatics Support", -->
<!--     Service.ID == "BCC-HS-2606" ~ "Bioinformatics Project Request", -->
<!--     TRUE ~ Service.Type -->
<!--   )) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- table(ilabs.bioinfo.simple$Service.Type) -->
<!-- ``` -->

<!-- The column names are long and somewhat confusing, simplify them with -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple <- ilabs.bioinfo.simple %>% -->
<!--   dplyr::mutate(Service.Type = case_when( -->
<!--     Service.Type == "Bioinformatics Project Request" ~ "Project", -->
<!--     Service.Type == "Prepaid Bioinformatics Support" ~ "Project_Prepaid", -->
<!--     TRUE ~ Service.Type  # Keep any other values unchanged -->
<!--   )) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- table(ilabs.bioinfo.simple$Service.Type) -->
<!-- ``` -->

<!-- The Service.Type "Teaching - Duan" is an attempt to capture a 1:1 -->
<!-- teaching event. The price is the same as project work and the category -->
<!-- is not consistently tracked in bioinformatics data. The line between -->
<!-- project work and 1:1 teaching is grey, move these values to project -->
<!-- work, these kinds of hours would often just be tracked by project hours -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple <- ilabs.bioinfo.simple %>% -->
<!--   dplyr::mutate(Service.Type = case_when( -->
<!--     Service.Type == "Teaching - Duan" ~ "Project", -->
<!--     TRUE ~ Service.Type  # Keep any other values unchanged -->
<!--   )) -->
<!-- ``` -->

<!-- Any values in here named with "Prepaid Bioinformatics Support (KI)" can -->
<!-- be changed to "Project_Prepaid" I checked this with the following -->
<!-- command: -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple %>% dplyr::filter(Service.Type == "Prepaid Bioinformatics Support (KI)") %>% dplyr::select(Service.ID) %>% table() -->
<!-- ``` -->

<!-- this repairs the data: -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple <- ilabs.bioinfo.simple %>% -->
<!--   dplyr::mutate(Service.Type = case_when( -->
<!--     Service.Type == "Prepaid Bioinformatics Support (KI)" ~ "Project_Prepaid", -->
<!--     TRUE ~ Service.Type  # Keep any other values unchanged -->
<!--   )) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- table(ilabs.bioinfo.simple$Service.Type) -->
<!-- ``` -->

<!-- ## Comparing Service.Type and Category -->

<!-- Two columns appear to contain similar data, Service.Type, Category This -->
<!-- commands explore the relationships between the values -->

<!-- ```{r} -->
<!-- table(ilabs.bioinfo.simple$Service.Type,ilabs.bioinfo.simple$Category) -->
<!-- ``` -->

<!-- Almost always, these are appropriately annotated Hours of Service but -->
<!-- there are questions. -->

<!-- These are likely to be an entry glitch that is catching the end of month -->
<!-- prepaid billing events for prepaid support. All of the other entries of -->
<!-- this type are under Service.Type Project_Prepaid and usually have -->
<!-- Unit.of.Measure of Month -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple %>% dplyr::filter(Category == "Prepaid Support" & ilabs.bioinfo.simple$Service.Type == "Project_Prepaid") %>%  -->
<!--   dplyr::select(Unit.of.Measure) %>% table() -->

<!-- ilabs.bioinfo.simple %>% dplyr::filter(Category == "Prepaid Support" & ilabs.bioinfo.simple$Service.Type == "Project") %>%  -->
<!--   dplyr::select(Unit.of.Measure) %>% table() -->
<!-- ``` -->

<!-- Check the each values in more detail: -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple %>% dplyr::filter(Category == "Prepaid Support" &  -->
<!--                                          ilabs.bioinfo.simple$Service.Type == "Project" & -->
<!--                                          ilabs.bioinfo.simple$Unit.of.Measure == "each") -->
<!-- ``` -->

<!-- Based on all this, rows where Category == "Prepaid Support" are billing -->
<!-- events for prepaid support that should be excluded from reporting remove -->
<!-- them with: -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple <- ilabs.bioinfo.simple %>%  -->
<!--   dplyr::filter(!Category %in% c("Prepaid Support")) -->
<!-- dim(ilabs.bioinfo.simple) -->
<!-- ``` -->

<!-- Recheck the relationships between Service.Type, Category -->

<!-- ```{r} -->
<!-- table(ilabs.bioinfo.simple$Service.Type,ilabs.bioinfo.simple$Category) -->
<!-- ``` -->

<!-- There is another imperfection in the data with Training, check what the -->
<!-- 123 look like. Note the trailing space in this category -->

<!-- these are mis-entered training events. The distinction is not important -->
<!-- a ilabs.bioinfo.simple\$Category -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple %>% dplyr::filter(Category == "Hours of Service " & Service.Type == "Training_Session") #%>%  -->
<!--   #dplyr::select() #%>% table() -->
<!-- ``` -->

<!-- these are inconsistently entered training events. The distinction is not -->
<!-- important because the information is captured in Service.Type and the -->
<!-- column Category is no longer necessary, remove it -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple <- ilabs.bioinfo.simple %>%  -->
<!--   dplyr::select(-Category) -->
<!-- dim(ilabs.bioinfo.simple) -->
<!-- ``` -->

<!-- ## Repairing Customer.Institute -->

<!-- The information in Customer.Institute is verbose and results in a -->
<!-- cluttered view, alter these values -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple <- ilabs.bioinfo.simple %>% -->
<!--   dplyr::mutate(Customer.Institute = case_when( -->
<!--     Customer.Institute == "Boston Children's Hospital" ~ "BCH", -->
<!--     Customer.Institute == "KI-External" ~ "KI.ext", -->
<!--     Customer.Institute == "Massachusetts Institute of Technology" ~ "MIT", -->
<!--     Customer.Institute == "Rutgers University" ~ "Rutgers", -->
<!--     TRUE ~ Customer.Institute  # Keep any other values unchanged -->
<!--   )) -->
<!-- ``` -->

<!-- ## Repairing Customer.Lab data -->

<!-- This column is important because it will be used to join faculty -->
<!-- annotation data (see faculty.groups) -->

<!-- ```{r} -->
<!-- table(ilabs.bioinfo.simple$Customer.Lab) -->
<!-- ``` -->

<!-- Parse the Customer.Lab data by removing the " Lab" string and extracting -->
<!-- the correct parenthetical value for one entry, the (Gio) nickname is -->
<!-- problematic, remove that -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple <- ilabs.bioinfo.simple %>% -->
<!--   mutate(Customer.Lab = case_when( -->
<!--     Customer.Lab == "Traverso, (Gio) Carlo Giovanni (Koch) Lab" ~ "Traverso, Gio (Koch)", -->
<!--     TRUE ~ Customer.Lab -->
<!--   )) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple <- ilabs.bioinfo.simple %>% -->
<!--   mutate( -->
<!--     temp.Inst = str_extract(Customer.Lab, "(?<=\\()(.+?)(?=\\))"),  # Extract text in () -->
<!--     Customer.Lab = str_remove(Customer.Lab, " \\([^)]+\\) Lab$")  # Remove " (value) Lab" from end -->
<!--   ) -->
<!-- ``` -->

<!-- Check the modifications -->

<!-- ```{r} -->
<!-- table(ilabs.bioinfo.simple$Customer.Lab) -->
<!-- ``` -->

<!-- compare the new tmp.Inst data to the Customer.Institute data -->

<!-- ```{r} -->
<!-- table(ilabs.bioinfo.simple$Customer.Institute,ilabs.bioinfo.simple$temp.Inst) -->
<!-- ``` -->

<!-- These data indicate that the parenthetical value in Customer.Lab is -->
<!-- sometimes mis-entered and of no value Customer.Institute is all that is -->
<!-- necessary, remove the temp.Inst column -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple <- ilabs.bioinfo.simple %>%  -->
<!--   dplyr::select(-temp.Inst) -->
<!-- dim(ilabs.bioinfo.simple) -->
<!-- ``` -->

<!-- ## Repair the Charge.Name column -->

<!-- This column contains incorrect and mixed information. First, there is -->
<!-- information in there about training session details. Copy this column to -->
<!-- a new column called "Training.Details" - that can then be modified. -->
<!-- There is a lot of inconsistency in the names -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple <- ilabs.bioinfo.simple %>% dplyr::mutate(Training.Detail = Charge.Name) -->
<!-- ``` -->

<!-- The repairs below are not quite perfect but almost there. it might be -->
<!-- helpful to associate which staff member teaches which session at some -->
<!-- point -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple <- ilabs.bioinfo.simple %>% -->
<!--   mutate(Training.Detail = case_when( -->
<!--     str_detect(Training.Detail, regex("project|teaching", ignore_case = TRUE)) ~ "", -->
<!--     TRUE ~ Training.Detail -->
<!--   )) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple <- ilabs.bioinfo.simple %>% -->
<!--   mutate(Training.Detail = str_remove(Training.Detail, "^Training Session: ")) -->

<!-- ilabs.bioinfo.simple <- ilabs.bioinfo.simple %>% -->
<!--   mutate(Training.Detail = str_remove(Training.Detail, "^Training Session:")) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple <- ilabs.bioinfo.simple %>% -->
<!--   dplyr::mutate(Training.Detail = str_replace_all(Training.Detail, "Hands-on", "Hands_on")) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple <- ilabs.bioinfo.simple %>% -->
<!--   dplyr::mutate(Training.Detail = str_replace_all(Training.Detail, "principles.", "principles - ")) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple <- ilabs.bioinfo.simple %>% -->
<!--   dplyr::mutate(Training.Detail = str_replace_all(Training.Detail, ",  ", " - ")) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple <- ilabs.bioinfo.simple %>% -->
<!--   mutate(Training.Detail = str_remove(Training.Detail, " *-.*$")) -->
<!-- ``` -->

<!-- Check the modifications -->

<!-- ```{r} -->
<!-- table(ilabs.bioinfo.simple$Training.Detail) -->
<!-- ``` -->

<!-- ## Charge.name edits -->

<!-- After extraction of the training informaion from Charge.Name, this -->
<!-- column really captures which staff member did the project work Change -->
<!-- it's -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple <- ilabs.bioinfo.simple %>% -->
<!--   mutate(Charge.Name = case_when( -->
<!--     str_detect(Charge.Name, regex("Training Session", ignore_case = TRUE)) ~ NA_character_, -->
<!--     TRUE ~ Charge.Name -->
<!--   )) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple <- ilabs.bioinfo.simple %>% -->
<!--   dplyr::mutate(Charge.Name = str_remove(Charge.Name, "Project *- *")) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple <- ilabs.bioinfo.simple %>% -->
<!--   dplyr::mutate(Charge.Name = str_remove(Charge.Name, "Teaching *- *")) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple <- ilabs.bioinfo.simple %>% -->
<!--   mutate(Charge.Name = str_replace(Charge.Name, "CO-OP \\(for sponsor projects only\\)", "CO-OP_sponsorProj")) -->
<!-- ``` -->

<!-- Check the modifications -->

<!-- ```{r} -->
<!-- table(ilabs.bioinfo.simple$Charge.Name) -->
<!-- ``` -->

<!-- ## working with Notes column -->

<!-- There is a desire to stratify the project summaries into 2 groups, Data -->
<!-- Management and Project -->

<!-- Create a new column called "Work.Detail" from the column Service.Type -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple <- ilabs.bioinfo.simple %>% dplyr::mutate(Work.Detail = Service.Type) -->
<!-- ``` -->

<!-- Identify the data management service IDs -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple %>% dplyr::filter(Charge.Name %in% c("CharlieD", "Taisha")) %>% dplyr::select(Service.ID) %>% table() -->
<!-- ``` -->

<!-- This identifies some problems with the data. Rename the Kamm mistate -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple <- ilabs.bioinfo.simple %>% -->
<!--   mutate(Service.ID = str_replace(Service.ID,  -->
<!--                                    "\\tBCC-KAMM prepaid bioinformatics 2023-0.2",  -->
<!--                                    "BCC-KAMM prepaid bioinformatics 2023-0.2")) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple %>% dplyr::filter(Charge.Name %in% c("CharlieD", "Taisha")) %>% dplyr::select(Service.ID) %>% table() -->
<!-- ``` -->

<!-- Modify the Work.Detail column so that it has values for Data_Management -->
<!-- This logic is potentially imperfect Data managment if: Charge.Name in -->
<!-- this list: CharlieD,CO-OP,CO-OP_sponsorProj,Nathan,Taisha it is -->
<!-- Service.ID in this list: BCC-ANK-prepaid -->
<!-- bioinformatics-0.2,BCC-ENGELWARD prepaid bioinformatics -->
<!-- 2022-0.45,BCC-ENGELWARD prepaid bioinformatics 2023-0.45, BCC-ENGELWARD -->
<!-- prepaid bioinformatics-0.425,BCC-KAMM prepaid bioinformatics -->
<!-- 2022-0.2,BCC-KAMM prepaid bioinformatics 2023-0.2,BCC-KAMM prepaid -->
<!-- bioinformatics-0.2, BCC-LGG-prepaid bioinformatics-0.2,BCC-WHITE prepaid -->
<!-- bioinformatics-0.2 -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple <- ilabs.bioinfo.simple %>% -->
<!--   mutate(Work.Detail = case_when( -->
<!--     Charge.Name %in% c("CharlieD", "CO-OP", "CO-OP_sponsorProj", "Nathan", "Taisha") | -->
<!--     Service.ID %in% c("BCC-ANK-prepaid bioinformatics-0.2", -->
<!--                       "BCC-ENGELWARD prepaid bioinformatics 2022-0.45", -->
<!--                       "BCC-ENGELWARD prepaid bioinformatics 2023-0.45", -->
<!--                       "BCC-ENGELWARD prepaid bioinformatics-0.425", -->
<!--                       "BCC-KAMM prepaid bioinformatics 2022-0.2", -->
<!--                       "BCC-KAMM prepaid bioinformatics 2023-0.2", -->
<!--                       "BCC-KAMM prepaid bioinformatics-0.2", -->
<!--                       "BCC-LGG-prepaid bioinformatics-0.2", -->
<!--                       "BCC-WHITE prepaid bioinformatics-0.2") ~ "Data_Management", -->
<!--     TRUE ~ Work.Detail -->
<!--   )) -->
<!-- ``` -->

<!-- The notes column captures many instances on Allen work in data managment -->
<!-- that is attributed to Huiming, repair this Identify Allen Notes for Data -->
<!-- Management -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple %>%  -->
<!--   dplyr::filter(Work.Detail == "Data_Management") %>% -->
<!--   dplyr::filter(str_detect(Notes, regex("Allen", ignore_case = TRUE))) %>% -->
<!--   dplyr::select(Notes) %>%  -->
<!--   table() -->
<!-- ``` -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple <- ilabs.bioinfo.simple %>% -->
<!--   mutate(Charge.Name = case_when( -->
<!--     Notes %in% c("Allen","Allen CSBC","Allen Impact","Allen Metnet","Allen SRP") ~ "Allen", -->
<!--     TRUE ~ Charge.Name -->
<!--   )) -->
<!-- ``` -->

<!-- Check the modifications -->

<!-- ```{r} -->
<!-- table(ilabs.bioinfo.simple$Charge.Name) -->
<!-- ``` -->

<!-- # Join Faculty Annotations -->

<!-- Add a Fac.Key column to the data by removing the " Lab" string and -->
<!-- extracting the correct parenthetical value for one entry, the (Gio) -->
<!-- nickname is problematic, remove that -->

<!-- ```{r} -->
<!-- faculty.groups <- faculty.groups %>% dplyr::mutate(Fac.Key = Row.Labels) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- faculty.groups <- faculty.groups %>% -->
<!--   mutate(Fac.Key = case_when( -->
<!--     Fac.Key == "Traverso, (Gio) Carlo Giovanni (Koch) Lab" ~ "Traverso, Gio (Koch)", -->
<!--     TRUE ~ Fac.Key -->
<!--   )) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- faculty.groups <- faculty.groups %>% -->
<!--   mutate( -->
<!--     temp.Inst = str_extract(Fac.Key, "(?<=\\()(.+?)(?=\\))"),  # Extract text in () -->
<!--     Fac.Key = str_remove(Fac.Key, " \\([^)]+\\) Lab$")  # Remove " (value) Lab" from end -->
<!--   ) -->
<!-- ``` -->

<!-- Check the Keys -->

<!-- ```{r} -->
<!-- # Check unique values in ilabs.bioinfo.simple$Customer.Lab -->
<!-- unique(ilabs.bioinfo.simple$Customer.Lab) -->

<!-- # Check unique values in faculty.groups$Fac.Key -->
<!-- unique(faculty.groups$Fac.Key) -->

<!-- # See which values match -->
<!-- intersect(ilabs.bioinfo.simple$Customer.Lab, faculty.groups$Fac.Key) -->

<!-- # See which don't match -->
<!-- setdiff(ilabs.bioinfo.simple$Customer.Lab, faculty.groups$Fac.Key) -->
<!-- ``` -->

<!-- there was a mismatched row, get rid of it -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple <- ilabs.bioinfo.simple %>% dplyr::filter(!is.na(Customer.Lab)) -->
<!-- ``` -->

<!-- ## Adding SF faculty annotations -->

<!-- Sarah provided annotations for color: -->

<!-- the colors are: -->

<!-- None = MIT non-sponsor+external -->

<!-- Blue = MIT sponsor -->

<!-- Green = KI member -->

<!-- Pink = KI affiliated (e.g., clinical investigators, prior members, -->
<!-- postdocs w/ funding, core staff) -->

<!-- These were edited for brevity and to avoid special characters and a -->
<!-- annotation tibble was created: -->

<!-- ```{r} -->
<!-- fac.annot <- tibble( -->
<!--   Color = c("none", "blue", "green", "pink"), -->
<!--   Fac.Group = c("MIT_non.sponsor.external", "MIT_sponsor", "KI_member", "KI_affiliate") -->
<!-- ) -->
<!-- ``` -->

<!-- Join the annotations to the faculty table -->

<!-- ```{r} -->
<!-- dim(faculty.groups) -->
<!-- faculty.groups <- faculty.groups %>% -->
<!--   merge(fac.annot,  -->
<!--         by.x = "Color",  -->
<!--         by.y = "Color",  -->
<!--         all.x = TRUE,  -->
<!--         all.y = TRUE) -->
<!-- dim(faculty.groups) -->
<!-- ``` -->

<!-- Join the faculty annotations -->

<!-- ```{r} -->
<!-- faculty.groups.toJoin <- faculty.groups %>% dplyr::select(Fac.Key,Fac.Group) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- dim(ilabs.bioinfo.simple) -->
<!-- ilabs.bioinfo.simple.final <- ilabs.bioinfo.simple %>% -->
<!--   merge(faculty.groups.toJoin,  -->
<!--         by.x = "Customer.Lab",  -->
<!--         by.y = "Fac.Key",  -->
<!--         all.x = TRUE,  -->
<!--         all.y = FALSE) -->
<!-- dim(ilabs.bioinfo.simple.final) -->
<!-- ``` -->

<!-- # Export the data -->

<!-- ```{r} -->
<!-- write.xlsx(ilabs.bioinfo.simple.final,"Parsed_Bioinfo_ilabs.xlsx") -->
<!-- ``` -->

<!-- # Visualization examples -->

<!-- ## Barcharts of data management -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple.final %>% -->
<!--   filter(Work.Detail == "Data_Management") %>% -->
<!--   group_by(Fac.Group, Charge.Name) %>% -->
<!--   summarise(total_quantity = sum(Quantity, na.rm = TRUE), .groups = 'drop') %>% -->
<!--   complete(Fac.Group, Charge.Name, fill = list(total_quantity = 0)) %>% -->
<!--   ggplot(aes(x = Fac.Group, y = total_quantity, fill = Charge.Name)) + -->
<!--   geom_col(position = "dodge") + -->
<!--   labs(title = "Data Management Total Quantity by Faculty Group and Charge Name", -->
<!--        x = "Faculty Group", -->
<!--        y = "Total Quantity", -->
<!--        fill = "Charge Name") + -->
<!--   theme_minimal() + -->
<!--   theme(axis.text.x = element_text(angle = 45, hjust = 1)) -->
<!-- ``` -->

<!-- ## Barcharts of project -->

<!-- ```{r} -->
<!-- ilabs.bioinfo.simple.final %>% -->
<!--   filter(Work.Detail %in% c("Project", "Project_Prepaid")) %>% -->
<!--   group_by(Fac.Group, Charge.Name) %>% -->
<!--   summarise(total_quantity = sum(Quantity, na.rm = TRUE), .groups = 'drop') %>% -->
<!--   complete(Fac.Group, Charge.Name, fill = list(total_quantity = 0)) %>% -->
<!--   ggplot(aes(x = Fac.Group, y = total_quantity, fill = Charge.Name)) + -->
<!--   geom_col(position = "dodge") + -->
<!--   labs(title = "Project Total Quantity by Faculty Group and Charge Name", -->
<!--        x = "Faculty Group", -->
<!--        y = "Total Quantity", -->
<!--        fill = "Charge Name") + -->
<!--   theme_minimal() + -->
<!--   theme(axis.text.x = element_text(angle = 45, hjust = 1)) -->
<!-- ``` -->

<!-- # Data Repair - BMC file -->

<!-- ## select relevant rows -->

<!-- ```{r} -->
<!-- bmc.ext.filt <- ext.billing.infStor %>% -->
<!--   dplyr::filter(Charge.Name %in% c("Informatics Support", -->
<!--                                    "Informatics Support - FEB", -->
<!--                                    "Informatics Support - JAN")) -->
<!-- ``` -->

<!-- ## Repair the date -->

<!-- The imported dates are in a machine-readable format that is not -->
<!-- valuable. convert to human readable: -->

<!-- ```{r} -->
<!-- bmc.ext.filt$Purchase.Date <- as.Date(bmc.ext.filt$DATE, origin = "1899-12-30") -->
<!-- ``` -->

<!-- remove the old date and other unnecssary columns -->

<!-- ```{r} -->
<!-- bmc.ext.filt <- bmc.ext.filt %>% dplyr::select(-COST,-DATE,-CostObj,-TargetObject,-`_ID`,-PROJID) -->
<!-- ``` -->

<!-- Add missing columns needed to match ilabs data -->

<!-- ```{r} -->
<!-- bmc.ext.filt <- bmc.ext.filt %>% dplyr::mutate(Service.ID = "") -->
<!-- bmc.ext.filt <- bmc.ext.filt %>% dplyr::mutate(Customer.Institute = "") -->
<!-- bmc.ext.filt <- bmc.ext.filt %>% dplyr::mutate(Charge.Name = "") -->
<!-- bmc.ext.filt <- bmc.ext.filt %>% dplyr::mutate(Notes = "FileMaker Data") -->
<!-- bmc.ext.filt <- bmc.ext.filt %>% dplyr::mutate(Unit.of.Measure = "hour") -->
<!-- bmc.ext.filt <- bmc.ext.filt %>% dplyr::mutate(Price.Type = "External non-profit") -->
<!-- bmc.ext.filt <- bmc.ext.filt %>% dplyr::mutate(Center = "MIT-BMC-CORE") -->
<!-- bmc.ext.filt <- bmc.ext.filt %>% dplyr::mutate(Training.Detail = "") -->
<!-- bmc.ext.filt <- bmc.ext.filt %>% dplyr::mutate(Work.Detail = "Project") -->
<!-- bmc.ext.filt <- bmc.ext.filt %>% dplyr::mutate(Fac.Group = "MIT_non.sponsor.external") -->
<!-- ``` -->

<!-- Rename columns -->

<!-- ```{r} -->
<!-- bmc.ext.filt <- bmc.ext.filt %>% dplyr::rename(Service.Type = Usage.TYPE) -->
<!-- ``` -->

<!-- Reorder columns to match ilabs -->

<!-- ```{r} -->
<!-- bmc.ext.filt <- bmc.ext.filt %>% dplyr::select(Customer.Lab,Service.ID,Service.Type,Customer.Name, -->
<!--                                                                        Customer.Institute,Charge.Name,Notes,Unit.of.Measure, -->
<!--                                                                        Quantity,Price.Type,Purchase.Date,Center,Training.Detail,Work.Detail,Fac.Group) -->
<!-- ``` -->

<!-- Write to file for manual edits -->

<!-- ```{r} -->
<!-- #write.xlsx(bmc.ext.filt,file="sl_external.Filt.xlsx") -->
<!-- ``` -->

<!-- Read in manually edited file -->

<!-- ```{r} -->
<!-- bmc.ext.filt.edits <- read.xlsx("sl_external.Filt.xlsx") -->
<!-- ``` -->

<!-- Bind ilabs and filemaker data to one file -->

<!-- ```{r} -->
<!-- ilabs.filemaker.final <- rbind(ilabs.bioinfo.simple.final,bmc.ext.filt.edits) -->
<!-- ``` -->

<!-- # write final output file -->

<!-- ```{r} -->
<!-- write.xlsx(ilabs.filemaker.final,file="BCC_ilabs.filemaker.final.xlsx") -->
<!-- ``` -->

# write session info

Capturing information about the R session helps document your work

```{r}
sessionInfo()
writeLines(capture.output(sessionInfo()), "dataCleaning_sessionInfo.txt")
```
