---
title: "IGT - Data Cleaning, Summary, and Visualizations"
author: "Charlie Whittaker"
output: 
  html_document:
    toc: true
    toc_depth: 4
  html_notebook: default
date: "2/20/2026"
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = TRUE,
  warning = TRUE,
  error = TRUE,
  cache = TRUE,
  cache.lazy = FALSE,
  fig.retina = 2
)

# For bioinformatics (Seurat, parallel)
options(future.globals.maxSize = 8e10)

# Optional: seed for reproducibility
set.seed(42)
```

```{r}
library(openxlsx)
library(tidyverse)
library(ggrepel)
library(ggpubr)
library(purrr)
library(knitr)
library(lubridate)
library(patchwork)
```

# Data Import

## Read in ilabs stock data and select Biopolymers rows

Select both cores with MIT BioMicro Center in name

```{r}
ilabs <- read.xlsx("../data/stock_ilabs.xlsx")
ilabs.igt <- ilabs %>% dplyr::filter(Core.Name %in% c("KI Genomics Core / MIT BioMicro Center","KI Genomics Core / MIT BioMicro Center and KI High Throughput Sciences Core"))
rm(ilabs)
```

### Repair ilabs date

```{r}
ilabs.igt$Purchase.Date <- as.Date(ilabs.igt$Purchase.Date, origin = "1899-12-30")
```

### Check year table to make sure the 2 cores are just a rename - they are, rename started in 2024

```{r}
ilabs.igt %>%
  dplyr::mutate(Year = lubridate::year(Purchase.Date)) %>%
  dplyr::count(Year, Core.Name) %>%
  tidyr::pivot_wider(names_from = Year, values_from = n, values_fill = 0)
```

## Read in bmc-external

```{r}
bmc.ext <- read.xlsx("../data/ExtUsage2025_2510.xlsx")
```

### Data Repair - bmc external

#### simplify the imported dataframe

This first step identifies empty or near-empty columns and summarizes
the results. The bmc.ext data as none of these low complexity columns.

```{r}
# simple.cols <- map_df(names(bmc.ext), ~{
#   col <- bmc.ext[[.]]
#   n_unique <- n_distinct(col[!is.na(col)])
# 
#   tibble(
#     column = .,
#     value = if_else(all(is.na(col)),
#                     "empty",
#                     if_else(n_unique == 1,
#                            as.character(unique(col[!is.na(col)])[1]),
#                            NA_character_))
#   )
# }) %>%
#   filter(!is.na(value))
# 
# kable(simple.cols, format = "simple")
```

If the data frame had these columns, they could be excluded with this code. 

```{r}
# cat("Dimensions before simple column exclusion:", dim(bmc.ext), "\n")
# bmc.ext <- bmc.ext %>% dplyr::select(-all_of(simple.cols$column))
# cat("Dimensions after simple column exclusion:", dim(bmc.ext), "\n")
```

#### Repair date

```{r}
bmc.ext$Purchase.Date <- as.Date(bmc.ext$DATE, origin = "1899-12-30")
```

#### Exclude DSMI-related charges from bmc.ext

```{r}
bmc.ext.igt <- bmc.ext %>% dplyr::filter(!TYPE %in% c("Bioinformatics", "Computing"))
```

#### Compare column names in ilabs.dsmi and bmc.ext

```{r}
colnames(ilabs.igt)
colnames(bmc.ext.igt)
```

#### Rename columns to match ilabs

```{r}
bmc.ext.igt <- bmc.ext.igt %>% dplyr::rename(Customer.Name = USER)
bmc.ext.igt <- bmc.ext.igt %>% dplyr::mutate(Service.ID = `_ID`)
bmc.ext.igt <- bmc.ext.igt %>% dplyr::rename(Customer.Lab = LAB)
bmc.ext.igt <- bmc.ext.igt %>% dplyr::rename(Charge.Name = DETAIL)
bmc.ext.igt <- bmc.ext.igt %>% dplyr::rename(Quantity = NUM)
bmc.ext.igt <- bmc.ext.igt %>% dplyr::rename(Total.Price = COST)
bmc.ext.igt <- bmc.ext.igt %>% dplyr::rename(Category = TYPE)
```

#### Add missing columns needed to match ilabs data

```{r}
bmc.ext.igt <- bmc.ext.igt %>% dplyr::mutate(Price = round(Total.Price/Quantity,2))
bmc.ext.igt <- bmc.ext.igt %>% dplyr::mutate(Service.Type = "ToUpdate")
bmc.ext.igt <- bmc.ext.igt %>% dplyr::mutate(Customer.Department = "")
bmc.ext.igt <- bmc.ext.igt %>% dplyr::mutate(Customer.Institute = "")
bmc.ext.igt <- bmc.ext.igt %>% dplyr::mutate(Unit.of.Measure = "ToUpdate")
bmc.ext.igt <- bmc.ext.igt %>% dplyr::mutate(Price.Type = "")
bmc.ext.igt <- bmc.ext.igt %>% dplyr::mutate(Core.Name = "bmc.ext")
bmc.ext.igt <- bmc.ext.igt %>% dplyr::mutate(Center = "")
bmc.ext.igt <- bmc.ext.igt %>% dplyr::mutate(Usage.Type = "")
bmc.ext.igt <- bmc.ext.igt %>% dplyr::mutate(Notes = "")
```

#### Select ilabs columns from bmc.ext.igt

```{r}
bmc.ext.igt.to_rbind <- bmc.ext.igt %>%
  dplyr::select(Service.ID, Service.Type, Customer.Name, Customer.Lab, Customer.Department, 
                Customer.Institute, Charge.Name, Notes, Unit.of.Measure, Quantity, Price, 
                Total.Price, Price.Type, Purchase.Date, Core.Name, Center, Category, 
                Usage.Type)
```

#### Check the column names

```{r}
colnames(ilabs.igt)
colnames(bmc.ext.igt.to_rbind)
```

### Join ilabs and bmc.ext.igt

```{r}
igt.final <- rbind(ilabs.igt,bmc.ext.igt.to_rbind)
```

### Exclude HTS rows using Category information

```{r}
igt.final <- igt.final %>% dplyr::filter(!stringr::str_starts(Category, "HTS"))
```

### Check dates before reporting range selection

```{r, fig.width=12,fig.height=5}
igt.final %>%
  mutate(Year_Month = floor_date(Purchase.Date, "month")) %>%
  count(Year_Month) %>%
  arrange(Year_Month) %>%
  ggplot(aes(x = Year_Month, y = n)) +
  geom_col(fill = "steelblue") +
  scale_x_date(date_labels = "%b %Y", date_breaks = "1 month") +
  labs(title = "Records by Month",
       x = "Month",
       y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Select reporting date range

```{r}
cat("Dimensions of igt.final before reporting range selection:", dim(igt.final), "\n")
igt.final <- igt.final %>% dplyr::filter(between(Purchase.Date, as.Date("2020-05-01"), as.Date("2025-10-31")))
cat("Dimensions of igt.final after reporting range selection:", dim(igt.final), "\n")
```

### Check dates after reporting range selection

```{r, fig.width=12,fig.height=5}
igt.final %>%
  mutate(Year_Month = floor_date(Purchase.Date, "month")) %>%
  count(Year_Month) %>%
  arrange(Year_Month) %>%
  ggplot(aes(x = Year_Month, y = n)) +
  geom_col(fill = "steelblue") +
  scale_x_date(date_labels = "%b %Y", date_breaks = "1 month") +
  labs(title = "Records by Month",
       x = "Month",
       y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

# Faculty Annotation Work

### Import faculty key data

```{r}
fac.key <- read.xlsx("../data/FacultyKeys.xlsx")
```

### Associate Faculty.Key with igt.final

```{r}
igt.final <- igt.final %>%
  left_join(fac.key %>%
              select(ValueInData, Faculty.Key) %>%
              distinct(), 
            by = c("Customer.Lab" = "ValueInData"))
```

### Import sponsor annotation data

```{r}
fac.annot <- read.xlsx("../data/annotated.Faculty.Keys.xlsx")
```

### Join to igt.final using Faculty.Key

```{r}
igt.final <- igt.final %>%
  dplyr::left_join(
    fac.annot %>% dplyr::select(Faculty.Key, IGT_DSMI_Sponsor),
    by = "Faculty.Key"
  )
```

### Simplify sponsor associations for plotting

#### Import simplified categories

```{r}
sponsor.toPlot <- read.xlsx("../data/sponsor.Keys.ForPlotting.xlsx")
sponsor.toPlot
```

#### Join simplified sponsor data to igt.final

```{r}
igt.final <- igt.final %>%
  dplyr::left_join(sponsor.toPlot, by = c("IGT_DSMI_Sponsor" = "Sponsor"))
```

# Service Annotation

This may be necessary to facilitate data type selections for plotting

## Gather categrories and output data

Information about what was done is stored in 2 columns: Service.Type and Charge.Name

```{r}
# igt.final <- igt.final %>%
#   mutate(Service.Group.working = case_when(
#     Core.Name == "Biopolymers & Proteomics Core" ~ Service.Type,
#     Core.Name == "metpro" ~ Charge.Name,
#     TRUE ~ NA_character_  # default for any other cases
#   ))
# 
# cat("Dimensions of Service.Group.working dataframe:", dim(as.data.frame(table(igt.final$Service.Group.working))), "\n")
# 
# msbp.Service.Group.working.df <- as.data.frame(table(igt.final$Service.Group.working))
# write.xlsx(msbp.Service.Group.working.df,file="msbp.Service.Group.working.df.xlsx")
```

## Import manually annotated service groups

Manually annotate Service.Group based on data in Service.Group.working and import the results

```{r}
# msbp.Service.Groups <- read.xlsx("../data/msbp.Service.Group.Annotation.xlsx")
```

## Join Service.Group to igt.final

```{r}
# igt.final <- igt.final %>%
#   dplyr::left_join(
#     msbp.Service.Groups %>% dplyr::select(Service.Group.working, Service.Group),
#     by = "Service.Group.working"
#   )
```

# Data Visualizations

## Define Colors

```{r}
# Define colors for each Sponsor (adjust these hex codes as needed)
sponsor_colors <- c(
  "Center_Member" = "steelblue",
  "MIT_Sponsor" = "lightpink",
  "Non_Sponsor" = "orange"
)

# Define order of sponsors for stacking (top to bottom)
sponsor_order <- c("Center_Member", "MIT_Sponsor", "Non_Sponsor")
```

### Instrument Use: Assisted or Training

Include usage.type that includes "assisted", "training" or "unattended"
Include usage.type = N/A and Category = "SEM/FIB/Airyscan charges"
Include usage.type = N/A and Category = "Technical Time" and Charge.Name contains "Training",
Include usage.type = N/A and Category = TEM/Cryo-TEM charges and Charge.Name does NOT include "Sample" or "Cryoplunging"


```{r}
train.1 <- ilabs.nano %>% dplyr::filter(str_detect(Usage.Type, "Assisted|Training|Unattended"))
cat("Rows train1:", dim(train.1)[1], "\n")
train.2 <- ilabs.nano %>% dplyr::filter(Usage.Type == "N/A" & Category == "SEM/FIB/Airyscan charges")
cat("Rows train2:", dim(train.2)[1], "\n")
train.3 <- ilabs.nano %>% dplyr::filter(Usage.Type == "N/A" & Category == "Technical Time" & str_detect(Charge.Name, "Training"))
cat("Rows train3:", dim(train.3)[1], "\n")
train.4 <- ilabs.nano %>% dplyr::filter(Usage.Type == "N/A" & 
                                          Category == "TEM/Cryo-TEM charges" & 
                                          !str_detect(Charge.Name, "Cryoplunging") & 
                                          !str_detect(Charge.Name, regex("Sample",ignore_case = TRUE)))
cat("Rows train4:", dim(train.4)[1], "\n")
```

#### Combine the train objects

```{r}
train <- rbind(train.1,train.2,train.3,train.4)
```

### Prepare Instrument Use: Assisted or Training pie chart

```{r}
train.pie <- train %>%
  mutate(Sponsor_ForPlot = factor(Sponsor_ForPlot, levels = sponsor_order)) %>%
  group_by(Sponsor_ForPlot) %>%
  summarise(Quantity = sum(Quantity), 
            Num_Labs = if_else(first(Sponsor_ForPlot) == "Center_Member",
                               n_distinct(Faculty.Key[NanoFIB_Sponsor == "KI_member"]),
                               n_distinct(Faculty.Key)),
            .groups = "drop") %>%
  mutate(Percent = round((Quantity / sum(Quantity)) * 100, 0),
         Lab_Label = if_else(Sponsor_ForPlot == "Center_Member",
                             paste0(Num_Labs, " Members"),
                             paste0(Num_Labs, " Users")),
         Label = paste0(Quantity, "\n", Percent, "%\n", Lab_Label)) %>%
  ggplot(aes(x = "", y = Quantity, fill = Sponsor_ForPlot)) +
  geom_col() +
  geom_text(aes(label = Label), 
            position = position_stack(vjust = 0.5),
            size = 3,
            fontface = "bold",
            color = "white") +
  coord_polar("y", start = 0) +
  scale_fill_manual(values = sponsor_colors) +
  labs(title = "Instrument Use: Assisted or Training Hours",
       fill = "Sponsor") +
  theme_void() +
  theme(legend.position = "right")

train.pie
```

# Plot the pie charts figure

```{r,fig.width=16,fig.height=4}
trained.use.pie + train.pie + sample.pie + plot_layout(ncol = 3, guides = "collect")

combo.pie <- trained.use.pie + train.pie + sample.pie + plot_layout(ncol = 3, guides = "collect")
ggsave("figures/IGT_combo.pie.pdf", 
       plot = combo.pie, 
       dpi = 150, 
       width = 16, 
       height = 4, 
       units = "in")
```

# Write annotated output file

```{r}
write.xlsx(igt.final,"igt.final.xlsx")
```

# write session info

Capturing information about the R session helps document your work

```{r}
sessionInfo()
writeLines(capture.output(sessionInfo()), "IGT_processing_sessionInfo.txt")
```
