---
title: "ilabs Data Import and Pre-Processing"
author: "Charlie Whittaker"
output: 
  html_document:
    toc: true
    toc_depth: 4
  html_notebook: default
date: "12/20/2025"
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = TRUE,
  warning = TRUE,
  error = TRUE,
  cache = TRUE,
  cache.lazy = FALSE,
  fig.retina = 2
)

# For bioinformatics (Seurat, parallel)
options(future.globals.maxSize = 8e10)

# Optional: seed for reproducibility
set.seed(42)

# Clear knitr cache
unlink("hts_Processing_cache", recursive = TRUE)
```

```{r}
library(openxlsx)
library(tidyverse)
library(ggrepel)
library(ggpubr)
library(purrr)
library(knitr)
library(lubridate)
library(patchwork)
```

# Data Import

# Read in ilabs csv files

The files come in 2 parts because ilabs only allows 3 years worth of export at a time.
For this reporting, the date range is 1/1/20-10/31/25

Additional exports files can be added as needed

```{r}
Part1 <- read.csv("../data/01012020_12312022_ALL_Cores_Completed.csv")
Part2 <- read.csv("../data/01012023_10312025_ALL_Cores_Completed.csv")
```

## Join the parts together into a single ilabs R object

```{r}
ki.ilabs <- rbind(Part1,Part2)
```

```{r}
# Show ALL duplicate rows (including first occurrence)
tt <- ki.ilabs[duplicated(ki.ilabs, fromLast = FALSE) |
                duplicated(ki.ilabs, fromLast = TRUE), ]
dim(tt)
```

Check the dimensions of the assembled data

```{r}
cat("Dimensions of imported ilabs data before cleanup:", dim(ki.ilabs), "\n")
```

# Column Cleaning

## simplify the ilabs.bioinfo dataframe

The imported object has 47 columns, some columns have a single value,
some are empty. This first step identifies those columns and summarizes
the results

```{r}
simple.cols <- map_df(names(ki.ilabs), ~{
  col <- ki.ilabs[[.]]
  n_unique <- n_distinct(col[!is.na(col)])
  
  tibble(
    column = .,
    value = if_else(all(is.na(col)), 
                    "empty",
                    if_else(n_unique == 1,
                           as.character(unique(col[!is.na(col)])[1]),
                           NA_character_))
  )
}) %>%
  filter(!is.na(value))

kable(simple.cols, format = "simple")
```

Now that the information is captured, exclude those simple columns,
reported on above, and create a new simplified dataframe

```{r}
ki.ilabs.simple <- ki.ilabs %>% dplyr::select(-all_of(simple.cols$column))
cat("Dimensions of imported ilabs data after exluding low-content columns:", dim(ki.ilabs.simple), "\n")
```

```{r}
# Show ALL duplicate rows (including first occurrence)
tt <- ki.ilabs.simple[duplicated(ki.ilabs.simple, fromLast = FALSE) |
                duplicated(ki.ilabs.simple, fromLast = TRUE), ]
dim(tt)
```

## Select and Repair date data

This simplified dataframe has several columns that report dates but
Purchase.Date is the one that captures information about when the
service was provided. Exclude unnecessary dates and adjust the values to
so that they are human readable

```{r}
ki.ilabs.simple <- ki.ilabs.simple %>% dplyr::select(-Creation.Date,-Completion.Date,-Billing.Date,
                                                     -Date.file.sent.to.ERP,-Billing.Event.End.Date)
ki.ilabs.simple <- ki.ilabs.simple %>% mutate(Purchase.Date = parse_date_time(Purchase.Date, orders = "mdy HM"))
cat("Dimensions of imported ilabs data after exluding extra date columns:", dim(ki.ilabs.simple), "\n")
```

```{r}
# Show ALL duplicate rows (including first occurrence)
tt <- ki.ilabs.simple[duplicated(ki.ilabs.simple, fromLast = FALSE) |
                duplicated(ki.ilabs.simple, fromLast = TRUE), ]
dim(tt)
```

The year information can now be accessed with code like this:

```{r}
ki.ilabs.simple %>%
  mutate(year = year(Purchase.Date)) %>%
  group_by(year) %>%
  summarize(total_quant = sum(Quantity))
```

## email columns

For the purpose of CCSG reporting, email addresses are not needed.
customer name (Customer.Name) and lab (Customer.Lab) are captured in different columns
exclude these from the simplified data frame

```{r}
ki.ilabs.simple <- ki.ilabs.simple %>% dplyr::select(-User.Login.Email,-PI.Email,-Financial.Contact.Email)
cat("Dimensions of imported ilabs data after exluding unecessary email columns:", dim(ki.ilabs.simple), "\n")
```

```{r}
# Show ALL duplicate rows (including first occurrence)
tt <- ki.ilabs.simple[duplicated(ki.ilabs.simple, fromLast = FALSE) |
                duplicated(ki.ilabs.simple, fromLast = TRUE), ]
dim(tt)
```

## Additional columns to exlude for simplicity

some billing-related columns are not necessary also Asset.ID is not needed

```{r}
ki.ilabs.simple <- ki.ilabs.simple %>% dplyr::select(-Asset.ID,-Customer.Title,-Payment.Information,-Cost.Object,
                                                      -Billing.Status,-Invoice.Num,-Created.By)
cat("Dimensions of imported ilabs data after exluding other unnecessary columns:", dim(ki.ilabs.simple2), "\n")
```

## The column No.Charge.Justification are also probably not needed

Most of the very few annotations might still represent work and should still be counted in reporting.

```{r}
as.data.frame(table(ki.ilabs.simple$No.Charge.Justification))
```

a little more detail on these columns

```{r}
ki.ilabs.simple %>% dplyr::filter(No.Charge.Justification != "") %>% dplyr::select(Core.Name) %>% table()
ki.ilabs.simple %>% dplyr::filter(No.Charge.Justification != "") %>% dplyr::select(Quantity) %>% table()
```

Exclude the column

```{r}
ki.ilabs.simple <- ki.ilabs.simple %>% dplyr::select(-No.Charge.Justification)
cat("Dimensions of imported ilabs data after exluding No.Charge.Justification:", dim(ki.ilabs.simple), "\n")
```

```{r}
# Show ALL duplicate rows (including first occurrence)
tt <- ki.ilabs.simple[duplicated(ki.ilabs.simple, fromLast = FALSE) |
                duplicated(ki.ilabs.simple, fromLast = TRUE), ]
dim(tt)
```

# Additional Column Checks - this reports the number of unique values in the remaining 20 columns

```{r}
ki.ilabs.simple %>% summarize(across(everything(), n_distinct)) %>% as.data.frame() %>% t()
```

## Organization.Name is not helpful, information in captured in Customer.Institute

```{r}
table(ki.ilabs.simple$Organization.Name)
ki.ilabs.simple %>% dplyr::filter(Organization.Name == "Princeton Molecular Biology") %>% dplyr::select(Core.Name,Customer.Institute,Organization.Name)
```

Remove the column

```{r}
ki.ilabs.simple <- ki.ilabs.simple %>% dplyr::select(-Organization.Name)
cat("Dimensions of imported ilabs data after exluding Organization.Name:", dim(ki.ilabs.simple), "\n")
```

```{r}
# Show ALL duplicate rows (including first occurrence)
tt <- ki.ilabs.simple[duplicated(ki.ilabs.simple, fromLast = FALSE) |
                duplicated(ki.ilabs.simple, fromLast = TRUE), ]
dim(tt)
```

## The data in Core.ID is redundant with Core.Name

Check the relationship

```{r}
table(ki.ilabs.simple$Core.Name,ki.ilabs.simple$Core.ID)
```

Exclude Core.ID

```{r}
ki.ilabs.simple <- ki.ilabs.simple %>% dplyr::select(-Core.ID)
cat("Dimensions of imported ilabs data after exluding Code.ID:", dim(ki.ilabs.simple), "\n")
```

```{r}
# Show ALL duplicate rows (including first occurrence)
tt <- ki.ilabs.simple[duplicated(ki.ilabs.simple, fromLast = FALSE) |
                duplicated(ki.ilabs.simple, fromLast = TRUE), ]
dim(tt)
```

```{r}
ki.ilabs.simple %>% summarize(across(everything(), n_distinct)) %>% as.data.frame() %>% t()
```

All the remaining columns have some potential utility in report generation

# Export the ilabs data for use in other scripts

```{r}
write.xlsx(ki.ilabs.simple,"../data/stock_ilabs.xlsx")
```

# write session info for ilabs preprocessing

Capturing information about the R session helps document your work

```{r}
sessionInfo()
writeLines(capture.output(sessionInfo()), "ilabs_preprocessing_sessionInfo.txt")
```
