---
title: "TPSI - Data Cleaning, Summary, and Visualizations"
author: "Charlie Whittaker"
output: 
  html_document:
    toc: true
    toc_depth: 4
  html_notebook: default
date: "2/19/2026"
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = TRUE,
  warning = TRUE,
  error = TRUE,
  cache = TRUE,
  cache.lazy = FALSE,
  fig.retina = 2
)

# For bioinformatics (Seurat, parallel)
options(future.globals.maxSize = 8e10)

# Optional: seed for reproducibility
set.seed(42)
```

```{r}
library(openxlsx)
library(tidyverse)
library(ggrepel)
library(ggpubr)
library(purrr)
library(knitr)
library(lubridate)
library(patchwork)
```

# Data Import

## Read in ilabs stock data and select Biopolymers rows

```{r}
ilabs <- read.xlsx("../data/stock_ilabs.xlsx")
ilabs.tpsi <- ilabs %>% dplyr::filter(Core.Name %in% c("Microscopy Facility","Histology Core Facility"))
rm(ilabs)
```

### Repair ilabs date

```{r}
ilabs.tpsi$Purchase.Date <- as.Date(ilabs.tpsi$Purchase.Date, origin = "1899-12-30")
```

### Check dates before range selection

```{r, fig.width=12,fig.height=5}
ilabs.tpsi %>%
  mutate(Year_Month = floor_date(Purchase.Date, "month")) %>%
  count(Year_Month) %>%
  arrange(Year_Month) %>%
  ggplot(aes(x = Year_Month, y = n)) +
  geom_col(fill = "steelblue") +
  scale_x_date(date_labels = "%b %Y", date_breaks = "1 month") +
  labs(title = "Records by Month",
       x = "Month",
       y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Select reporting date range

```{r}
cat("Dimensions of ilabs.tpsi before reporting range selection:", dim(ilabs.tpsi), "\n")
ilabs.tpsi <- ilabs.tpsi %>% dplyr::filter(between(Purchase.Date, as.Date("2020-05-01"), as.Date("2025-10-31")))
cat("Dimensions of ilabs.tpsi after reporting range selection:", dim(ilabs.tpsi), "\n")
```

### Check dates after reporting range selection

```{r, fig.width=12,fig.height=5}
ilabs.tpsi %>%
  mutate(Year_Month = floor_date(Purchase.Date, "month")) %>%
  count(Year_Month) %>%
  arrange(Year_Month) %>%
  ggplot(aes(x = Year_Month, y = n)) +
  geom_col(fill = "steelblue") +
  scale_x_date(date_labels = "%b %Y", date_breaks = "1 month") +
  labs(title = "Records by Month",
       x = "Month",
       y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

# Faculty Annotation Work

### Import faculty key data

```{r}
fac.key <- read.xlsx("../data/FacultyKeys.xlsx")
```

### Associate Faculty.Key with ilabs.tpsi

```{r}
ilabs.tpsi <- ilabs.tpsi %>%
  left_join(fac.key %>%
              select(ValueInData, Faculty.Key) %>%
              distinct(), 
            by = c("Customer.Lab" = "ValueInData"))
```

### Import sponsor annotation data

```{r}
fac.annot <- read.xlsx("../data/annotated.Faculty.Keys.xlsx")
```

### Join sponsorship to ilabs.tpsi using Faculty.Key

```{r}
ilabs.tpsi <- ilabs.tpsi %>%
  dplyr::left_join(
    fac.annot %>% dplyr::select(Faculty.Key, Microscopy_Sponsor, Histology_Sponsor),
    by = "Faculty.Key"
  )
```

### Create an aggregated sponsor category for plotting

```{r}
ilabs.tpsi <- ilabs.tpsi %>%
  dplyr::mutate(Sponsor = case_when(
    Core.Name == "Histology Core Facility" ~ Histology_Sponsor,
    Core.Name == "Microscopy Facility" ~ Microscopy_Sponsor,
    TRUE ~ NA_character_
  ))
```


### Simplify sponsor associations for plotting

#### Import simplified categories

```{r}
sponsor.toPlot <- read.xlsx("../data/sponsor.Keys.ForPlotting.xlsx")
sponsor.toPlot
```

#### Join simplified sponsor data to ilabs.tpsi

```{r}
ilabs.tpsi <- ilabs.tpsi %>%
  dplyr::left_join(sponsor.toPlot, by = c("Sponsor" = "Sponsor"))
```

# Service Annotation

This may be necessary to facilitate data type selections for plotting

## Gather categrories and output data

Information about what was done is stored in 2 columns: Service.Type and Charge.Name

```{r}
# ilabs.tpsi <- ilabs.tpsi %>%
#   mutate(Service.Group.working = case_when(
#     Core.Name == "Biopolymers & Proteomics Core" ~ Service.Type,
#     Core.Name == "metpro" ~ Charge.Name,
#     TRUE ~ NA_character_  # default for any other cases
#   ))
# 
# cat("Dimensions of Service.Group.working dataframe:", dim(as.data.frame(table(ilabs.tpsi$Service.Group.working))), "\n")
# 
# msbp.Service.Group.working.df <- as.data.frame(table(ilabs.tpsi$Service.Group.working))
# write.xlsx(msbp.Service.Group.working.df,file="msbp.Service.Group.working.df.xlsx")
```

## Import manually annotated service groups

Manually annotate Service.Group based on data in Service.Group.working and import the results

```{r}
# msbp.Service.Groups <- read.xlsx("../data/msbp.Service.Group.Annotation.xlsx")
```

## Join Service.Group to ilabs.tpsi

```{r}
# ilabs.tpsi <- ilabs.tpsi %>%
#   dplyr::left_join(
#     msbp.Service.Groups %>% dplyr::select(Service.Group.working, Service.Group),
#     by = "Service.Group.working"
#   )
```

# Rate Modifiers

This is being paused now - using consistent units (each and hours instead)

## Import rate modifier data

```{r}
# rate.mod <- read.xlsx("../data/rateModifiers.xlsx")
# rate.mod
```

## Join the 2 different rate modifiers to ilabs.tpsi using SponsorCategory

This should be validated by looking for Customer.Lab values that have more than 1 rate.modifier

```{r}
# msbp.final <- msbp.final %>%
#   left_join(
#     rate.mod %>% select(SponsorCategory, BP_modifier),
#     by = c("BioPol_Sponsor" = "SponsorCategory")
#   ) %>%
#   left_join(
#     rate.mod %>% select(SponsorCategory, metpro_modifier),
#     by = c("MetPro_Sponsor" = "SponsorCategory")
#   ) %>%
#   mutate(rate.modifier = case_when(
#     Core.Name == "Biopolymers & Proteomics Core" ~ BP_modifier,
#     Core.Name == "metpro" ~ metpro_modifier,
#     TRUE ~ NA_real_
#   )) %>%
#   select(-BP_modifier, -metpro_modifier)  # clean up temporary columns
```

## Create adjusted Total.Price values based on assigned rate.modifier values

```{r}
# msbp.final <- msbp.final %>% dplyr::mutate(Adjusted.Total.Price = Total.Price * rate.modifier)
```

# Data Visualizations

## Define Colors

```{r}
# Define colors for each Sponsor (adjust these hex codes as needed)
sponsor_colors <- c(
  "Center_Member" = "steelblue",
  "MIT_Sponsor" = "lightpink",
  "Non_Sponsor" = "orange"
)

# Define order of sponsors for stacking (top to bottom)
sponsor_order <- c("Center_Member", "MIT_Sponsor", "Non_Sponsor")
```

## Pie Charts

### Pathology Consult

```{r}
ilabs.tpsi %>% dplyr::filter(str_detect(Usage.Type, "Consult")) %>% dplyr::select(Usage.Type) %>% table()
ilabs.tpsi %>% dplyr::filter(Usage.Type == "N/A" & Unit.of.Measure == "hour") %>% dplyr::select(Usage.Type) %>% table()
pc.1 <- ilabs.tpsi %>% dplyr::filter(str_detect(Usage.Type, "Consult"))
pc.2 <- ilabs.tpsi %>% dplyr::filter(Usage.Type == "N/A" & Unit.of.Measure == "hour")
path <- rbind(pc.1,pc.2)
dim(path)
```

### Prepare Pathology Consult pie chart

```{r}
path.pie <- path %>%
  mutate(Sponsor_ForPlot = factor(Sponsor_ForPlot, levels = sponsor_order)) %>%
  group_by(Sponsor_ForPlot) %>%
  summarise(Quantity = sum(Quantity), 
            Num_Labs = if_else(first(Sponsor_ForPlot) == "Center_Member",
                               n_distinct(Faculty.Key[Sponsor == "KI_member"]),
                               n_distinct(Faculty.Key)),
            .groups = "drop") %>%
  mutate(Percent = round((Quantity / sum(Quantity)) * 100, 0),
         Lab_Label = if_else(Sponsor_ForPlot == "Center_Member",
                             paste0(Num_Labs, " Members"),
                             paste0(Num_Labs, " Users")),
         Label = paste0(Quantity, "\n", Percent, "%\n", Lab_Label)) %>%
  ggplot(aes(x = "", y = Quantity, fill = Sponsor_ForPlot)) +
  geom_col() +
  geom_text(aes(label = Label), 
            position = position_stack(vjust = 0.5),
            size = 3,
            fontface = "bold",
            color = "white") +
  coord_polar("y", start = 0) +
  scale_fill_manual(values = sponsor_colors) +
  labs(title = "Pathology Consultation Hours",
       fill = "Sponsor") +
  theme_void() +
  theme(legend.position = "right")

path.pie
```

### Instrument Use: Assisted and Training

```{r}
instr <- ilabs.tpsi %>% dplyr::filter(Usage.Type %in% c("Assisted", 
                                                        "Assisted Imaging- Visium", 
                                                        "Assisted Use", 
                                                        "New User Training", 
                                                        "New User Training (remote)", 
                                                        "Training", 
                                                        "Unattended scan time"))
```

#### Check the data

```{r}
table(instr$Usage.Type)
```

### Prepare Instrument Use: Assisted and Training pie chart

```{r}
instr.pie <- instr %>%
  mutate(Sponsor_ForPlot = factor(Sponsor_ForPlot, levels = sponsor_order)) %>%
  group_by(Sponsor_ForPlot) %>%
  summarise(Quantity = sum(Quantity), 
            Num_Labs = if_else(first(Sponsor_ForPlot) == "Center_Member",
                               n_distinct(Faculty.Key[Sponsor == "KI_member"]),
                               n_distinct(Faculty.Key)),
            .groups = "drop") %>%
  mutate(Percent = round((Quantity / sum(Quantity)) * 100, 0),
         Lab_Label = if_else(Sponsor_ForPlot == "Center_Member",
                             paste0(Num_Labs, " Members"),
                             paste0(Num_Labs, " Users")),
         Label = paste0(Quantity, "\n", Percent, "%\n", Lab_Label)) %>%
  ggplot(aes(x = "", y = Quantity, fill = Sponsor_ForPlot)) +
  geom_col() +
  geom_text(aes(label = Label), 
            position = position_stack(vjust = 0.5),
            size = 3,
            fontface = "bold",
            color = "white") +
  coord_polar("y", start = 0) +
  scale_fill_manual(values = sponsor_colors) +
  labs(title = "Instrument Use: Assisted and Training Hours",
       fill = "Sponsor") +
  theme_void() +
  theme(legend.position = "right")

instr.pie
```

### Instrument Use: Trained

```{r}
inst.train <- ilabs.tpsi %>% dplyr::filter(Usage.Type %in% c("Beginner User", 
                                                             "Beginner Users", 
                                                             "Independent Use", 
                                                             "Independent Use: off-peak time", 
                                                             "Independent Use: peak time", 
                                                             "Independent User: off-peak time", 
                                                             "Independent User: peak time", "Super User"))
```

#### Check the data

```{r}
table(inst.train$Usage.Type)
```

### Prepare Instrument Use: Assisted and Training pie chart

```{r}
inst.train.pie <- inst.train %>%
  mutate(Sponsor_ForPlot = factor(Sponsor_ForPlot, levels = sponsor_order)) %>%
  group_by(Sponsor_ForPlot) %>%
  summarise(Quantity = sum(Quantity), 
            Num_Labs = if_else(first(Sponsor_ForPlot) == "Center_Member",
                               n_distinct(Faculty.Key[Sponsor == "KI_member"]),
                               n_distinct(Faculty.Key)),
            .groups = "drop") %>%
  mutate(Percent = round((Quantity / sum(Quantity)) * 100, 0),
         Lab_Label = if_else(Sponsor_ForPlot == "Center_Member",
                             paste0(Num_Labs, " Members"),
                             paste0(Num_Labs, " Users")),
         Label = paste0(Quantity, "\n", Percent, "%\n", Lab_Label)) %>%
  ggplot(aes(x = "", y = Quantity, fill = Sponsor_ForPlot)) +
  geom_col() +
  geom_text(aes(label = Label), 
            position = position_stack(vjust = 0.5),
            size = 3,
            fontface = "bold",
            color = "white") +
  coord_polar("y", start = 0) +
  scale_fill_manual(values = sponsor_colors) +
  labs(title = "Instrument Use: Trained Hours",
       fill = "Sponsor") +
  theme_void() +
  theme(legend.position = "right")

inst.train.pie
```

### Tissue Processing

```{r}
tiss <- ilabs.tpsi %>% dplyr::filter(Usage.Type == "N/A") %>% dplyr::filter(Unit.of.Measure %in% c("hour", "Sample", "set", "slide", "each"))
```

#### Check the data

```{r}
table(tiss$Unit.of.Measure)
```

#### Prepare Tissue Processing pie chart

```{r}
tiss.pie <- tiss %>%
    mutate(Sponsor_ForPlot = factor(Sponsor_ForPlot, levels = sponsor_order)) %>%
  group_by(Sponsor_ForPlot) %>%
  summarise(Quantity = sum(Quantity), 
            Num_Labs = if_else(first(Sponsor_ForPlot) == "Center_Member",
                               n_distinct(Faculty.Key[Sponsor == "KI_member"]),
                               n_distinct(Faculty.Key)),
            .groups = "drop") %>%
  mutate(Percent = round((Quantity / sum(Quantity)) * 100, 0),
         Lab_Label = if_else(Sponsor_ForPlot == "Center_Member",
                             paste0(Num_Labs, " Members"),
                             paste0(Num_Labs, " Users")),
         Label = paste0(Quantity, "\n", Percent, "%\n", Lab_Label)) %>%
  ggplot(aes(x = "", y = Quantity, fill = Sponsor_ForPlot)) +
  geom_col() +
  geom_text(aes(label = Label), 
            position = position_stack(vjust = 0.5),
            size = 3,
            fontface = "bold",
            color = "white") +
  coord_polar("y", start = 0) +
  scale_fill_manual(values = sponsor_colors) +
  labs(title = "Tissue Processing Units",
       fill = "Sponsor") +
  theme_void() +
  theme(legend.position = "right")

tiss.pie
```

# Plot the pie charts figure

```{r,fig.width=16,fig.height=4}
instr.pie + inst.train.pie + path.pie + tiss.pie + plot_layout(ncol = 4, guides = "collect")

combo.pie <- instr.pie + inst.train.pie + path.pie + tiss.pie + plot_layout(ncol = 4, guides = "collect")
ggsave("figures/TPSI_combo.pie.pdf", 
       plot = combo.pie, 
       dpi = 150, 
       width = 20, 
       height = 4, 
       units = "in")
```


# Write annotated output file

```{r}
write.xlsx(ilabs.tpsi,"tpsi.final.xlsx")
```

# write session info

Capturing information about the R session helps document your work

```{r}
sessionInfo()
writeLines(capture.output(sessionInfo()), "TPSI_processing_sessionInfo.txt")
```
